---
title: "StatR 503 Final Project"
author: "Johanna Krogh"
date: "Due June 22, 2017"
output: html_document
---

```{r echo=FALSE, warning = FALSE, message = FALSE}
setwd("C:/Users/Jo/Dropbox/UW StatR Cert/Spring 17")
library(ggplot2)
library(rpart)
library(magrittr)
library(randomForest)
library(doParallel)
source("mycv2.r")
source("classificationMetrics.r")



```

#Introduction:
The Mushroom Data Set was donated to the UCI Machine Learning Repository almost exactly 30 years ago. I am interested in using this dataset because I would like to get a better understanding of ho wto work with categorical data, and how to interpret the results from the various classification methods I will use. I want to see if there is a reliable method to identify poisonous mushrooms. My hypothesis, based on some descriptions of the dataset, is that there may be a reliable way but not an parsimonious way. Meaning, knowing to look for a 2-3 distinguishing elements of a mushroom isn't a way to determine if it's edible or poisonous.

#Materials and Methods:
As stated above, I found this dataset on the UCI Machine Learning Repository. The data was initially published in the 1981 edition of the Audubon Society Field Guide to North American Mushrooms. Before pre-processing, there are 8,124 instances. For my analysis, I'll split the dataset into train and test data. I'll use the same train and test datasets for each of my classification methods: CART and Random Forests. 

There is no missing data from this dataset. There are 22 attributes, all categorical. The full list of attributes is as such:
1. Cap-Shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s

2. Cap-Surface: fibrous=f,grooves=g,scaly=y,smooth=s

3. Cap-Color: brown=n,buff=b,cinnamon=c,gray=g,green=r, pink=p,purple=u,red=e,white=w,yellow=y 

4. Bruises?: bruises=t,no=f

5. Odor: almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s

6. Gill-Attachment: attached=a,descending=d,free=f,notched=n

7. Gill-Spacing: close=c,crowded=w,distant=d

8. Gill-Size: broad=b,narrow=n

9. Gill-Color: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e, white=w,yellow=y

10. Stalk-Shape: enlarging=e,tapering=t

11. Stalk-Root: bulbous=b,club=c,cup=u,equal=e, rhizomorphs=z,rooted=r,missing=?

12. Stalk-Surface-Above-Ring: fibrous=f,scaly=y,silky=k,smooth=s

13. Stalk-Surface-Below-Ring: fibrous=f,scaly=y,silky=k,smooth=s

14. Stalk-Color-Above-Ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y

15. Stalk-Color-Below-Ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y

16. Veil-Type: partial=p,universal=u

17. Veil-Color: brown=n,orange=o,white=w,yellow=y

18. Ring-Number: none=n,one=o,two=t

19. Ring-Type: cobwebby=c,evanescent=e,flaring=f,large=l, none=n,pendant=p,sheathing=s,zone=z

20. Spore-Print-Color: black=k,brown=n,buff=b,chocolate=h,green=r, orange=o,purple=u,white=w,yellow=y

21. Population: abundant=a,clustered=c,numerous=n, scattered=s,several=v,solitary=y

22. Habitat: grasses=g,leaves=l,meadows=m,paths=p, urban=u,waste=w,woods=d

```{r echo=FALSE}
mushroom <-  read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", header = TRUE, sep = ",")
names(mushroom) <- c("class", "cap.shape","cap.surface", "cap.color","bruises", "odor", "gill.attach", "gill.spacing", "gill.size", "gill.color", "stalk.shape", "stalk.root","stalk.surf.above.ring","stalk.surf.below.ring", "stalk.color.above.ring", "stalk.color.below.ring", "veil.type", "veil.color","ring.number", "ring.type", "spore.print.color", "population", "habitat")

```

##Explore the data:
From exploring the data, can we get a sense if the two classes, edible and poisonous, are clearly defined?

###Plot 1: 
If you find a mushroom in the woods, leaves, paths, or urban environment, and there tends to be several of them together, look out. It's unclear if you will be dealing with a poisonous or edible mushroom. 
```{r echo=FALSE}
ggplot(mushroom, aes(x=cap.color , y=population, color = class)) + geom_point(alpha=0.8) +geom_jitter(alpha=0.8) + facet_grid(.~habitat) + xlab("Cap Color") + ylab("Population") + ggtitle("Cap Color vs Population faceted by Habitat")
```

###Plot 2:
Mushrooms with one ring and spore print colors of black and brown, have a lot of overlap with poisonous and edibles classes!
```{r echo=FALSE}
ggplot(mushroom, aes(x=spore.print.color , y=population, color = class)) + geom_point(alpha=0.8) +geom_jitter(alpha=0.8) + facet_grid(.~ring.number) + xlab("Ring Number") + ylab("Population") + ggtitle("Spore Print Color vs Population faceted by Ring Number")
```

###Plot 3:
Some clearer grouping here. If you find mushrooms in the wood and they have no odor, be wary.
```{r echo=FALSE}
ggplot(mushroom, aes(x=population , y=odor, color = class)) + geom_point(alpha=0.8) +geom_jitter(alpha=0.8) + facet_grid(.~habitat) + xlab("Population") + ylab("Odor") + ggtitle("Population vs Odor faceted by Habitat")

```

#Results

##Classification and Regression Trees: CART
First we start with the CART method, the simple method. I split the training data into 10 CV groups, randomly. Then I tuned the CP values. I did a simple CART with the default CP value of 0.01. Then I did another simple CART but with the chosen CP value from our tuning


```{r}
Mcv <- sample(10, 6498, replace = TRUE) #CV groups for training data
trainid <- sample(8123, 6498, replace = FALSE)#train id data
Mtest <- mushroom[-trainid,]#test data
Mtrain <- mushroom[trainid,]#train data

```


###Simple CART Confusion Matrix
Plotting 'tree' doesn't show anything helpful. Prediction seems to be ok. There were eleven poisonous mushrooms that were classified as edible. Let's see if tuning CP gets us a better result.

Sensitivity: `r 843/(843+12)`
Specificity: `r 770/770`
```{r}
tree <- rpart(class~., data = Mtrain)
table(mushroom$class[-trainid], predict(tree,newdata=Mtest,type='class'))


```


###Tuning Cp
```{r}
tuneCps <- c(1e-5, 1e-4, seq(0.001, 0.01, 0.001),
             seq(0.02, 0.1, 0.01), 0.2)

n <- dim(Mtrain)[1] 
k <- length(tuneCps)
Mtree1 <- matrix(NA,nrow=n,ncol=k)

for (a in 1:k)  # Looping over the Cp values
    Mtree1[,a] <- mycv2(class~.,data=Mtrain, FUN=rpart,cvid=Mcv,cp=tuneCps[a],predtype='class')
```

###Results of tuning CP
Based on these results I've chosen to go with a CP value of 0.001. Also below is the plotted CV curve for CP tuning. Flat Error Rate jumps sharping after CP=0.02.
```{r}
Mtree1 %>% apply(2,function(x,ref) mean(x!=ref),ref=as.integer(Mtrain$class)) %>% round(3) 
```


```{r echo=FALSE}
plot(tuneCps, Mtree1 %>% apply(2,function(x,ref) mean(x!=ref),ref=as.integer(Mtrain$class)) %>% round(3), main="CART CV Curve",  xlab = "Cp", ylab = "Flat Error Rate", type = "l")

```


```{r echo=FALSE}
fe <- Mtree1 %>% apply(2, function(x,ref) mean(x!=ref), ref=as.integer(Mtrain$class))
bestfe <- which.min(fe)

```

###Predicting with best CP value
Prediction is slightly better. Now only 4 poisonous mushrooms were classified as edible.
Sensitivity:`r 843/(843+3)`
Specificity:`r 779/779`
```{r}
Mpred1 <- rpart(class~.,data=Mtest,control=list(cp=0.001))
table(Mtest$class, predict(Mpred1, type = "class"))
```
###Plot CART
I don't know how much I trust this tree. If a mushroom doesn't have an odor, that means it's poisonous? This tree leads me to believe there is a parsimonious way to determine whether or not a mushroom is edible or poisonous. 
```{r}
plot(Mpred1)
text(Mpred1)
```


##Random Forest
Now onto Random Forest. I enjoyed the 2D tuning we did in class, so after tuning the B parameter I tuned mtry & nodesize. 

###Tuning B
```{r}
bseq=100*c(seq(2,10,2),15,20,30,50) 
j=length(bseq)
MBtune=matrix(NA,nrow=dim(Mtrain)[1],ncol=j)
system.time({
for (a in 1:j)
{
 set.seed(321)
 MforB<-randomForest(class~.,data=Mtrain,ntree=bseq[a])
 MBtune[,a]<-MforB$predicted
}})

```

###Results from tuning B
These results are troubling. They are all 0. Does this mean there is an issue with my data or my code? Or does this just mean that choosing the most parmimonious B value is straight forward? I cannot find fault with the data or code, so I will move forward with a chosen B value of 200.
```{r}
bseq
MBtune %>% apply(2,flatErr,ref=MBtune[,j]) %>% round(3)
```

###Tuning RandomForests
Not only do I tune for mtry and nodesize, I also thought it would be fun to look at the weightedScore. 
```{r}
ftune <- expand.grid(m=c(2:6,8,10),detail=c(1:4,6,8)) 
ncore=3
cl <- makeCluster(ncore, type = "SOCK")
registerDoParallel(cl)
system.time(
tune1results<-foreach(a=1:dim(ftune)[1],.combine='rbind',.packages='randomForest')
 %dopar% {
 set.seed(414155) 
 tmp <- randomForest(class~.,data=Mtrain,ntree=1500, mtry=ftune$m[a],
 nodesize=ftune$detail[a])
 c(flatErr(tmp$predicted,Mtrain$class),weightedScore (tmp$predicted,Mtrain$class, c(1, 10)))
 })
stopCluster(cl)
```

```{r echo=FALSE}
ftune$flat <- tune1results[,1] 
ftune$inv <- tune1results[,2]
```

###Results from tuning
Let's just look at the first six settings. There over 40 for each. The winning score is m = 2, with one node. 
```{r}
head(ftune[ftune$flat==min(ftune$flat),])
head(ftune[ftune$inv==min(ftune$inv),])
```

###Predict in Random Forest
With our winning  settings, let's predict! The results are almost too good to be true - perfect prediction!
```{r}
MRF<-randomForest(class~.,data=Mtrain,ntree=200,mtry=2,nodesize=1)
predRF <- predict(MRF,newdata=Mtest)

table(Mtest$class,predRF)
```

#Discussion:
Clearly, Random Forest is the best option for predicting what class a mushroom will be. However, the simple CART method, even without tuning CP, wasn't that bad! When it comes to determining if a food will kill you or not, it's better to be safe than sorry. As I stated in my hypothesis, I thought there would be a reliable but not a parsimonious way to determine the class. I still don't totally trust that CART tree we were able to plot above. A little research online shows my instinct to be good. There are some suggestions for how to pick out edible from poisonous, but overall the best advice is: don't eat a mushroom unless you are 100% sure of what it is. 
