---
title: "StatR 503 HW4"
author: "Johanna Krogh"
date: "Due May 22, 2017"
output: html_document
---

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(e1071)
library(mosaic)
library(boot)
```

##Problem 1: Characterizing Player Heights
```{r }
NBA <- read.csv("http://faculty.washington.edu/eliezg/teaching/StatR503/data/NBAcensus.csv")

```

###A. Basic Stats
Mean = `r mean(NBA$Height)`
Variance = `r var(NBA$Height)`
Skewness = `r skewness(NBA$Height)`
```{r}
hist(NBA$Height, xlab = "Height in Inches", main = "Histogram of NBA Heights")

```

###B. Skewness Bootstrap

```{r message=FALSE, warning=FALSE}

skew.boot <- function(d, i){
  skewness(d[i])
}
Height.bs <- boot(NBA$Height, skew.boot, R=1000, sim = "ordinary")

Height.CI <- apply(Height.bs$t,2,quantile, p=c(0.025, 0.5, 0.975))

hist(Height.bs$t, xlab="Height Skewness", main="Bootstrap Skewness of Height")
abline(v=c(Height.CI[1], Height.CI[3]), col = "red" )

```

95% Confidence Interval: (`r Height.CI[1]`, `r Height.CI[3]`)

##Problem 2: Fitting Distributions

###A.Log-Likelihoods
```{r message=FALSE, warning = FALSE}
LL.normal <- function(param, x){
    mu <- param[1]
  sigma <- param[2]
  -sum(log(dnorm(x,mu,sigma) ) )
}

LL.weibull <- function(param, x){
    lambda <- param[1]
  beta <- param[2]
  -sum(log(dweibull(x,lambda,beta) ) )
}

```

###B. Maximum Likelihood Estimates

#### Fit for normal
```{r message=FALSE, warning=FALSE}
p0 <- c(78, 3)
Like.fitN <- optim(p0, LL.normal, x=NBA$Height)
Like.fitN$par
```

#### Fit for Weibull
```{r}
p1 <- c(1, 79)
Like.fitW <- optim(p1, LL.weibull, x=NBA$Height)
Like.fitW$par
```

###C. Visualization and model selection

The Weibull distribution curve fits the histogram much better. The shape parameter allows the curve to skew left or right. Our log likelihood function gave us a parameter of shape = 27, modeling the left skew. This makes sense for our data set because we are looking at NBA heights, which will be skewed left compared to the rest of the population.

What criterion is most important for comparing the two models? It's difficult for me to wrap my head around the Weibull parameters. For this dataset I would say the shape of the curve is the most important.
```{r}
NBAseq <- seq(69,87, length.out=100) #NBAseq helps space the points out equally
hist(NBA$Height, xlab="Height in Inches", main="NBA Height", freq=FALSE) 

points(NBAseq,dnorm(NBAseq, mean = 79.119634, sd =3.428596 ), type = "l", col="blue") #plot normal distribution with params
points(NBAseq, dweibull(NBAseq,27.43594, 80.70287), type ="l", col = "green") #plot weibull distribution with params
legend("topright", c("Normal", "Weibull"),col=c("blue", "green"), lty=1)

```

###D. Confidence Intervals
The CI's are very similar. The bootstrap method is slight shifted more positive.

####Hessian for Normal
```{r message=FALSE, warning=FALSE}
param.fit.N <- optim(p0, LL.normal, x=NBA$Height, hessian=TRUE)#estimate for Fisher information matrix. Inverse of FI gives covariance for parameters
Sigma.N <- solve(param.fit.N$hessian)#Sigma used to note variance matrix
se.N <- sqrt(diag(Sigma.N))
cbind(hat = param.fit.N$par, CI.low = param.fit.N$par  - 1.96*se.N, CI.high = param.fit.N$par  + 1.96*se.N)

```

####Hessian for Weibull
```{r message=FALSE, warning=FALSE}
param.fit.W <- optim(p1, LL.weibull, x=NBA$Height, hessian=TRUE)
Sigma.W <- solve(param.fit.W$hessian)
se.W <- sqrt(diag(Sigma.W))
cbind(hat = param.fit.W$par, CI.low = param.fit.W$par  - 1.96*se.W, CI.high = param.fit.W$par  + 1.96*se.W)

```

####Bootstrap CI Normal
```{r message=FALSE, warning=FALSE}
MLE.boot.N <- function(d, i){
  Like.fitN2 <- optim(p0, LL.normal, x=d[i])  
  Like.fitN2$par
}
Height.MLE.N <- boot(NBA$Height, MLE.boot.N, R=1000, sim = "ordinary")

Height.CI.N <- apply(Height.MLE.N$t,2,quantile, p=c(0.025, 0.975))
Height.CI.N
```

####Bootstrap CI Weibull
```{r message=FALSE, warning=FALSE}
MLE.boot.W <- function(d, i){
  Like.fitW2 <- optim(p1, LL.weibull, x=d[i])  
  Like.fitW2$par
}
Height.MLE.W <- boot(NBA$Height, MLE.boot.W, R=1000, sim = "ordinary")

Height.CI.W <- apply(Height.MLE.W$t,2,quantile, p=c(0.025, 0.975))
Height.CI.N
```

